---
title: "Code to analyse data for Metagenomic Complexity study"
author: "Mun Hua Tan"
output:
  pdf_document: default
  html_document: default
---

# Date: 27/05/2025
# Paper: Metagenomic complexity of high, seasonal transmission of Plasmodium spp. in asymptomatic carriers in Northern Sahelian Ghana.
# Description: This code uses files from "data" folder to generate main figures in the paper

# ENV: Load packages & create general lists/data frames
```{r message=FALSE}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(ggsci)
library(mgcv)
library(sf)
library(wesanderson)
library(ggrepel)

ups_list <- list("upsA", "non-upsA")
AgeGroups_list <- list("6-10 Years", "11-20 Years", "21-39 Years", "40+ Years")
WBvol_list <- list("1uL", "10uL", "50uL", "100uL")

WBvol_pairs_df <- data.frame(pairs=c("1uL & 10uL", "1uL & 50uL", "1uL & 100uL",
                                     "10uL & 50uL", "10uL & 100uL", "50uL & 100uL"),
                             volS=c("1uL", "1uL", "1uL", "10uL", "10uL", "50uL"),
                             volB=c("10uL", "50uL", "100uL", "50uL", "100uL", "100uL"),
                             volS_label=c("WB1", "WB1", "WB1", "WB10", "WB10", "WB50"),
                             volB_label=c("WB10", "WB50", "WB100", "WB50", "WB100", "WB100"))

##Repertoire size threshold
repsize_thres_nonzero <- 1 # 1 DBLa type
repsize_thres_popgen <- 20 # 20 DBLa types

## Function: read data tables
read_table_in <- function(x) {
  x <- read.table(x, sep="\t", header=TRUE, check.names=FALSE)
} 
```

## Unzip some files
```{bash}
unzip data/gadm41_GHA_shp.zip -d data/
gunzip -c1 data/S8_DBS_pRBC_Repeats_OtuTable.txt.gz > data/S8_DBS_pRBC_Repeats_OtuTable.txt
```

## Figure 2
```{r}
## Load input file
isolates_field_pRBC <- read_table_in("data/S8_pRBC_Field_IsolateMetadata.txt")

## Arrange isolates for heatmap ordering: sorting by parasite density, then species detections by volume
isolates_field_pRBC_orderIsolates <- isolates_field_pRBC %>%
  select(Isolate, Source, NumTypes, ParasiteuL, starts_with("DBS_"), starts_with("WB100_")) %>%
  pivot_wider(names_from=Source, values_from=NumTypes) %>%
  arrange(desc(ParasiteuL),
          desc(DBS_Pf), desc(WB100_Pf),
          desc(DBS_Pm), desc(WB100_Pm),
          desc(DBS_Po), desc(WB100_Po),
          desc(DBS_Pv), desc(WB100_Pv))

## Set factor levels
isolates_field_pRBC <- isolates_field_pRBC %>% 
  mutate(Source=factor(Source, levels=WBvol_list),
         Isolate=factor(Isolate, levels=isolates_field_pRBC_orderIsolates$Isolate),
         AgeGrp=factor(AgeGrp, levels=AgeGroups_list))

## Create long-form labels for Age Groups with counts
xlabs <- isolates_field_pRBC %>%
  select(Isolate, AgeGrp) %>%
  distinct() %>%
  count(AgeGrp) %>%
  mutate(label_long = paste0(AgeGrp, "\n(n=", n, ")"))

## Join long labels to main data
isolates_field_pRBC <- isolates_field_pRBC %>%
  left_join(xlabs, by="AgeGrp") %>%
  mutate(label_long=factor(label_long, levels=xlabs$label_long)) # Set factor levels

## Prepare data for tile plot of species detection
plot_tile_data <- isolates_field_pRBC %>%
  select(Isolate, starts_with("DBS_"), starts_with("WB100_"), AgeGrp, label_long) %>%
  distinct() %>%
  pivot_longer(-c(Isolate, AgeGrp, label_long), names_to="Source", values_to="value") %>%
  mutate(value=ifelse(value == 0, NA, value),
         Source=str_replace(Source, "WB100_", "100uL_"),
         Source=str_replace(Source, "DBS_", "DBS_")) %>%
  separate(Source, into=c("Source", "Species"), sep="_") %>%
  mutate(Source=factor(Source, levels=c("DBS", "100uL")),
         Isolate=factor(Isolate, levels=isolates_field_pRBC_orderIsolates$Isolate)) %>%
  filter(Species != "Pv")

## Create tile plot
plot_tile_All_species <- ggplot(plot_tile_data, aes(x=Source, y=Isolate, fill=Species)) +
  geom_tile(color="white", aes(alpha=value)) +
  scale_fill_manual(name="Species", values=c("#548B54", "#79CDCD", "#8B3E2F", "black")) +
  scale_alpha(range=c(1, 1), na.value=0) +
  scale_x_discrete(name="Source") +
  scale_y_discrete(name="Isolate") +
  facet_grid(label_long~Species, space="free", scales="free") +
  theme_bw() +
  theme(panel.grid=element_blank(),
        axis.title=element_text(size=11),
        axis.text.x=element_text(size=11),
        axis.text.y=element_blank(),
        legend.position="none",
        strip.text=element_text(size = 11))

## Figure 2
# Save as image (tiff)
ggsave("Figure2.tiff", plot_tile_All_species,
       device="tiff", dpi=300, width=4, height=6)

# Display
plot_tile_All_species
```

## Figure 3, panel a
```{r}
## Load input file
isolates_field_pRBC <- read_table_in("data/S8_pRBC_Field_IsolateMetadata.txt")
isolates_field_DBS  <- read_table_in("data/S8_DBS_Field_IsolateMetadata.txt")

## Arrange isolates for heatmap ordering: sorting by parasite density, then species detections by volume
isolates_field_pRBC_orderIsolates <- isolates_field_pRBC %>%
  select(Isolate, Source, NumTypes, ParasiteuL, starts_with("DBS_"), starts_with("WB100_")) %>%
  pivot_wider(names_from=Source, values_from=NumTypes) %>%
  arrange(desc(ParasiteuL),
          desc(DBS_Pf), desc(WB100_Pf),
          desc(DBS_Pm), desc(WB100_Pm),
          desc(DBS_Po), desc(WB100_Po),
          desc(DBS_Pv), desc(WB100_Pv))

## Extract MOI per species for all species
species_list <- c("Pf", "Pm", "Poc", "Pow")

file_isolates_MOI <- map_dfr(species_list, function(sp) {
  moi_col <- paste0(sp, "MOI")
    
  # WB data
  field_data <- isolates_field_pRBC %>%
    select(Isolate, Source, AgeGrp, NumTypes, MOI=.data[[moi_col]]) %>%
    mutate(Species=sp)
  
  # DBS data
  dbs_data <- isolates_field_DBS %>%
    filter(Isolate %in% isolates_field_pRBC_orderIsolates$Isolate) %>%
    select(Isolate, Source, AgeGrp, NumTypes, MOI=.data[[moi_col]]) %>%
    mutate(Species=sp)
  
  bind_rows(field_data, dbs_data)
})

## Include only MOI > 0 (i.e. remove MOI = 0)
file_isolates_MOI <- file_isolates_MOI %>% filter(MOI > 0)

## Set factor levels
file_isolates_MOI <- file_isolates_MOI %>%
  mutate(Source=factor(Source, levels=c("DBS", WBvol_list)),
         Isolate=factor(Isolate, levels=isolates_field_pRBC_orderIsolates$Isolate),
         AgeGrp=factor(AgeGrp, levels=AgeGroups_list))

## Create long-form labels for Age Groups with counts for P. falciparum
xlabs <- file_isolates_MOI %>%
  filter(Species == "Pf" & Source %in% WBvol_list) %>%
  select(Isolate, AgeGrp) %>%
  distinct() %>%
  count(AgeGrp) %>%
  mutate(label_long=paste0(AgeGrp, "\n(n=", n, ")"))

## Join long labels to main data
file_isolates_MOI <- file_isolates_MOI %>%
  left_join(xlabs, by="AgeGrp") %>%
  mutate(label_long=factor(label_long, levels=xlabs$label_long)) # Set factor levels

## Get max MOI for scale limits
max_MOI <- max(file_isolates_MOI$MOI, na.rm = TRUE)

## Create MOI tile plot (WB, for only Pf)
plot_tile_MOI_WB <- file_isolates_MOI %>%
  filter(Species == "Pf", Source %in% WBvol_list) %>%
  ggplot(., aes(x=Source, y=Isolate, fill=as.numeric(MOI))) +
  geom_tile() +
  scale_fill_viridis_c(name="Pf-MOI", option="inferno", na.value="white",
                       limits=c(1, max_MOI), breaks=seq(1, max_MOI, 1),
                       values=c(0, 0.2, 1), direction=-1, begin=0.15, end=1.0) +
  scale_x_discrete(name="Volume") +
  scale_y_discrete(name="Isolate") +
  facet_grid(label_long~"pRBC", space="free", scale="free") +
  theme_bw() +
  theme(panel.grid=element_blank(),
    axis.text.y=element_blank(),
    strip.text=element_text(size=12),
    legend.key.height=unit(3.5, "cm"),
    legend.position="right")

## Create MOI tile plot (DBS, for only Pf)
# Get isolate-label mapping (before filtering)
isolate_labels <- file_isolates_MOI %>%
  filter(Species == "Pf") %>%
  select(Isolate, label_long) %>%
  distinct()

# Get complete only Isolate and Source, then join back label_long
file_isolates_MOI_DBS <- file_isolates_MOI %>%
  filter(Species == "Pf", Source == "DBS") %>%
  select(Isolate, Source, MOI) %>%
  complete(Isolate=levels(file_isolates_MOI$Isolate), Source="DBS") %>%
  inner_join(isolate_labels, by="Isolate") %>%
  mutate(Source=factor(Source, levels=c("DBS", WBvol_list)),
         Isolate=factor(Isolate, levels=levels(file_isolates_MOI$Isolate)),
         label_long=factor(label_long, levels=levels(file_isolates_MOI$label_long)))

plot_tile_MOI_DBS <- ggplot(file_isolates_MOI_DBS, aes(x=Source, y=Isolate, fill=as.numeric(MOI))) +
  geom_tile() +
  scale_fill_viridis_c(name="Pf-MOI", option="inferno", na.value="grey80",
                       limits=c(1, max_MOI), breaks=seq(1, max_MOI, 1),
                       values=c(0, 0.2, 1), direction=-1, begin=0.15, end=1.0) +
  scale_x_discrete(name="Source") +
  scale_y_discrete(name="Isolate") +
  facet_grid(label_long~"DBS", space="free", scale="free") +
  theme_bw() +
  theme(panel.grid=element_blank(),
        axis.text.y=element_blank(),
        axis.title.x=element_blank(),
        legend.position="none",
        strip.text.x=element_text(size=12),
        strip.text.y=element_blank())

## Combine WB and DBS plots
plot_tile_MOI_combine <- ggarrange(plot_tile_MOI_DBS, plot_tile_MOI_WB, nrow=1, align="h", widths=c(0.25, 1))

## Figure 3a
# Display
plot_tile_MOI_combine
```

## Figure 3, panel b
```{r}
## Load input file
isolates_field_pRBC <- read_table_in("data/S8_pRBC_Field_IsolateMetadata.txt")

## Include only isolates with found_min1_code Y-Y-Y-Y
found_min_code <- paste0("found_min", repsize_thres_nonzero, "_code")
isolates_field_pRBC <- isolates_field_pRBC %>% filter(.data[[found_min_code]] == "Y-Y-Y-Y")
  
## Extract info
isolates_field_pRBC <- isolates_field_pRBC %>%
  select(Isolate, Source, PfMOI) %>%
  pivot_longer(-c(Isolate, Source), names_to="metric") %>%
  mutate(metric=gsub("PfMOI", "Pf-MOI", metric)) %>%
  mutate(Source=factor(Source, levels=WBvol_list))
  
## Statistical test
kw_pval_all <- data.frame()

pw_metric <- "Pf-MOI"
pw_metric_name <- "MOI"
  
file_isolates_field_metric <- isolates_field_pRBC %>% filter(metric == pw_metric)
   
# Shapiro test for normality
shapiro_norm <- file_isolates_field_metric %>% shapiro_test(value)

# Statistical test (Friedman test)
friedman_nstats <- file_isolates_field_metric %>% friedman_test(value ~ Source | Isolate)

# Statistical test (Wilcoxon paired test with holm correction)
kw_pval <- file_isolates_field_metric %>%
  pairwise_wilcox_test(value ~ Source,
                       paired=TRUE,
                       p.adjust.method="holm",
                       alternative="two.sided",
                       comparisons=list(c("1uL", "10uL"), c("1uL", "50uL"), c("1uL", "100uL"),
                                        c("10uL", "50uL"), c("10uL", "100uL"), c("50uL", "100uL"))) %>% 
  add_xy_position(x="Source", step.increase=0.05) %>%
  mutate(metric=pw_metric) %>%
  mutate(alternative="none")
  
# Save
kw_pval_all <- rbind(kw_pval_all, kw_pval)

## Plot boxplots
## Create long-form labels
xlabs <- file_isolates_field_metric %>%
  select(Isolate, Source) %>%
  distinct() %>%
  count(Source) %>%
  mutate(label_long=paste0(Source, "\n(n=", n, ")"))

# Create boxplots
plot_paired_boxplot <- file_isolates_field_metric %>%
  ggboxplot(., x="Source", y="value", id="Isolate",
           line.color="gray", fill="Source",  line.size=0.4, line.type=NA, add="jitter") +
  scale_fill_jco(name="Volume") +
  stat_pvalue_manual(kw_pval, xmin="group1", xmax="group2", label="p.adj.signif", hide.ns=TRUE, tip.length=0, vjust=0.9) +
  labs(subtitle = get_test_label(friedman_nstats, detailed=TRUE)) +
  theme_bw() +
  scale_x_discrete(name="Volume", labels=xlabs$label_long) +
  scale_y_continuous(name=pw_metric, breaks=scales::pretty_breaks(10), limits=c(0,NA)) +
  theme(legend.text=element_text(size=12),
        panel.grid=element_blank(),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14)) +
  facet_wrap(metric~., scales="free") +
  theme(strip.text=element_text(size=12))

## Figure 3b
# Display
plot_paired_boxplot
```

## Figure 3, panel c
```{r}
## Load input file
isolates_diff_FD <- read_table_in("data/id96.FIELD.fold_diff.txt")

## Filter for 1ul & 100uL comparisons specific to Pf-MOI
isolates_diff_FD <- isolates_diff_FD %>% filter(label == "1uL & 100uL" & diff_metric == "Pf-MOI")

## GAM model
# Fit GAM model with maximum k based on unique sample volume values
k_max <- n_distinct(isolates_diff_FD$volS_value)
gam_model <- gam(diff_value ~ s(volS_value, k=k_max), data=isolates_diff_FD, method="REML")

# Add GAM predictions to the data
isolates_diff_FD <- isolates_diff_FD %>% mutate(gam_pred=predict(gam_model, newdata=.))

## Create scatter plot
# Prepare data for plotting only 1uL vs 100uL
plot_data <- isolates_diff_FD %>%
  filter(diff_metric == "Pf-MOI" & volS_name == "1uL" & volB_name == "100uL") %>%
  group_by(volS_name, volB_name, volS_value, diff_value, label_long) %>%
  summarise(pointsize = n(), .groups = "drop")

# Create scatter plot
plot_scatter_FD_MOI_1vs100_model <- ggplot(plot_data, aes(x=volS_value, y=diff_value)) +
  geom_hline(yintercept=2^0, linetype="dashed", color="red", size=0.5) +
  geom_point(aes(fill=volS_name, size=pointsize), color="black", shape=21) +
  geom_line(data=isolates_diff_FD, aes(x=volS_value, y=gam_pred), color="darkgreen", size=0.7) +
  scale_x_continuous(name=expression(Pf-MOI[S]), breaks=scales::pretty_breaks(10), limits=c(0, NA)) +
  scale_y_continuous(name="Fold Difference", trans=scales::log2_trans(),
                     breaks=scales::trans_breaks("log2", function(x) 2^x),
                     labels=scales::trans_format("log2", scales::math_format(2^.x))) +
  scale_fill_jco(name=expression(Vol[S])) +
  scale_size(name="Number of\nisolates") +
  facet_wrap(~label_long, nrow = 2) +
  theme_bw() +
  theme(panel.grid= element_blank(),
        legend.position="right",
        legend.text=element_text(size=12),
        axis.text= element_text(size=12),
        axis.title=element_text(size=14),
        axis.text.x=element_text(angle=90, vjust=0.5),
        strip.text=element_text(size=12))

## Figure 3c
# Display
plot_scatter_FD_MOI_1vs100_model

```

## Figure 3, panel d
```{r}
## Load input file
isolates_field_pRBC <- read_table_in("data/S8_pRBC_Field_IsolateMetadata.txt")
otutable <- read_table_in("data/S8_DBS_pRBC_Repeats_OtuTable.txt")

## Filter isolates with >= 20 DBLa types
isolates_field_pRBC <- isolates_field_pRBC %>% filter(NumTypes >= repsize_thres_popgen)

## Summarise number of isolates per blood volume
isolates_field_pRBC_summary <- isolates_field_pRBC %>%
  count(Source, name="num_isolates") %>%
  mutate(AgeGrp="Total") %>%
  relocate(., AgeGrp, .before=num_isolates)

## Summarise DBLa richness per blood volume
info_DBLa_richness <- map_dfr(WBvol_list, function(vol) {
  
  # Filter isolates for current volume
  isolates_field_pRBC_vol <- isolates_field_pRBC %>% 
    filter(Source == vol)
  
  # Subset OTU table and calculate DBLa frequencies
  otutable_vol <- otutable %>%
    select("DBLa_type":"upsA_nonA", all_of(isolates_field_pRBC_vol$IsolateID)) %>%
    mutate(dbla_freq = rowSums(across(all_of(isolates_field_pRBC_vol$IsolateID)))) %>%
    relocate(dbla_freq, .after = "upsA_nonA") %>%
    filter(dbla_freq > 0)
  
  # Summarise DBLa richness by upsA/nonA classification
  otutable_vol %>%
    count(upsA_nonA, name="num_dbla") %>%
    mutate(prop_dbla=num_dbla/sum(num_dbla),
           Source=vol)
})

## Set factors
isolates_field_pRBC_summary <- isolates_field_pRBC_summary %>%
  mutate(Source=factor(Source, levels=WBvol_list),
         AgeGrp=factor(AgeGrp, levels=c("Total", AgeGroups_list)))

info_DBLa_richness <- info_DBLa_richness %>%
  mutate(Source=factor(Source, levels=WBvol_list),
         upsA_nonA=factor(upsA_nonA, levels=ups_list))

## Create barplot and line plot for DBLa richness and number of isolates per blood volume
# Get coefficient to scale for secondary axis
coeff <- 50

# Plot number of isolates & DBLa richness
plot_numIsolates_richness <- ggplot(info_DBLa_richness, aes(x=Source, y=num_dbla)) +
  geom_bar(stat="identity") +
  geom_point(filter(isolates_field_pRBC_summary, AgeGrp=="Total"), mapping=aes(x=Source, y=num_isolates*coeff)) +
  geom_line(filter(isolates_field_pRBC_summary, AgeGrp=="Total"), mapping=aes(x=Source, y=num_isolates*coeff, group=1)) +
  scale_y_continuous(name=expression(paste("DBL", alpha, " type richness")),
                     sec.axis=sec_axis(~./coeff, name="Number of isolates"),
                     breaks=scales::pretty_breaks(10)) +
  scale_x_discrete(name="Volume") +
  theme_bw() +
  theme(legend.title=element_blank(),
        legend.text=element_text(size=12),
        panel.grid=element_blank(),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14))

## Figure 3d
# Display
plot_numIsolates_richness

```

## Figure 3, panel e
```{r}
## Load input file
isolates_field_pRBC <- read_table_in("data/S8_pRBC_Field_IsolateMetadata.txt")
field_PTS <- read_table_in("data/id96.FIELD_POP.PTS.txt")

## Filter isolates with >= 20 DBLa types
isolates_field_pRBC <- isolates_field_pRBC %>% filter(NumTypes >= repsize_thres_popgen)

## Set factor levels
isolates_field_pRBC$Source <- factor(isolates_field_pRBC$Source, levels=WBvol_list)
field_PTS$label <- factor(field_PTS$label, levels=WBvol_list)

## Create long-form labels for blood volume with counts
xlabs <- isolates_field_pRBC
xlabs <- paste(levels(xlabs$Source),"\n(n=",table(xlabs$Source),")",sep="")
  
## Statistical test (Kruskal-Wallis + Dunn multiple test with holm correction)
stats_pval_kw <- field_PTS %>%
  kruskal_test(PTS_score ~ label) %>%
  mutate(Test="Kruskal-Wallis")

stats_pval_dunn <- field_PTS %>%
  dunn_test(PTS_score ~ label, p.adjust.method="holm", detailed=TRUE) %>%
  mutate(Test="Dunn's test") %>%
  add_xy_position(x="label", step.increase=0.1)

## Get summary
field_PTS_summary <- field_PTS %>%
  group_by(label) %>%
  summarise(min=min(PTS_score),
            max=max(PTS_score),
            mean=mean(PTS_score),
            median=median(PTS_score),
            Q1=quantile(PTS_score, probs=0.25),
            Q3=quantile(PTS_score, probs=0.75),
            .groups = "drop")

## Plot PTS
plot_PTS_AB <- ggplot(field_PTS, aes(x=label, y=PTS_score)) +
  geom_violin(aes(fill=label)) +
  geom_boxplot(width=0.1) +
  stat_pvalue_manual(stats_pval_dunn, xmin="group1", xmax="group2", label="p.adj.signif", hide.ns=TRUE, tip.length=0, vjust=0.7) +
  # labs(fill="Volume") +
  scale_x_discrete(name="Volume", labels=xlabs) +
  scale_y_continuous(name="PTS", breaks=scales::pretty_breaks(10), limits=c(0,1)) +
  scale_fill_jco(name="") +
  theme_bw() +
  theme(legend.position="right",
        legend.text=element_text(size=12),
        panel.grid=element_blank(),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14))
  
## Figure 3e
# Display
plot_PTS_AB

```

## Figure 3 composite
```{r}
coeff <- 50
  
## PLOTS
plot1 <- plot_tile_MOI_combine
plot2 <- plot_paired_boxplot
plot3 <- plot_scatter_FD_MOI_1vs100_model
plot4 <- plot_numIsolates_richness
plot5 <- plot_PTS_AB

## PLOT LAYOUT
combine1 <- ggarrange(plot2, NA, plot3, ncol=1, labels=c("b", "", "c"), heights=c(1, 0.02, 0.8))
combine2 <- ggarrange(plot4, NA, plot5, nrow=1, labels=c("d", "", "e"), widths=c(0.9, 0.02, 1))
combine1_3 <- ggarrange(plot1, NA, combine1, nrow=1, widths=c(0.9, 0.02, 1), labels=c("a", "", ""))
combine_all <- ggarrange(combine1_3, NA, combine2, ncol=1, heights=c(1, 0.01, 0.5))

## Figure 3
# Save as image (tiff)
ggsave(paste0("Figure3.tiff"),
         combine_all,
         device="tiff", dpi=300,
         width=10, height=12)
# Display
combine_all
```

## Figure 4
```{r}
## Prepare data for isolates with WBvol = 100uL with MOI > 0
file_isolates_MOI_100 <- file_isolates_MOI %>%
  select(-label_long) %>%
  filter(Source == WBvol_list[4] & !is.na(MOI)) %>%
  group_by(Isolate) %>%
  mutate(sum_complexity=sum(MOI), # Sum MOI to get metagenomic complexity
         n_species=n()) %>% # Add count of species
  ungroup() %>%
  arrange(sum_complexity, n_species) %>% # Order by increasing number of species and complexity
  mutate(Isolate) %>%
  mutate(Isolate=factor(Isolate, levels=unique(Isolate)), # Set factor levels
         AgeGrp=factor(AgeGrp, levels=AgeGroups_list))

## Create long-form labels for Age Group with counts
xlabs <- file_isolates_MOI_100 %>%
  select(Isolate, AgeGrp) %>%
  distinct() %>%
  count(AgeGrp) %>%
  mutate(label_long = paste0(AgeGrp, "\n(n=", n, ")")) %>%
  select(-n)

## Join long-form labels to main data
file_isolates_MOI_100 <- file_isolates_MOI_100 %>%
  left_join(xlabs, by="AgeGrp") %>%
  mutate(label_long = factor(label_long, levels = xlabs$label_long)) # Set factor levels

## Create barplot - stacked barplot of raw MOI counts
plot_bar_MOI_stack <- ggplot(file_isolates_MOI_100, aes(x=Isolate, y=as.numeric(MOI), fill=Species)) +
  geom_bar(stat="identity", position="stack") +
  scale_x_discrete(name="Isolate") +
  scale_y_continuous("Metagenomic Complexity") +
  scale_fill_manual(name="Species", values=c("#548B54", "#79CDCD", "#8B3E2F", "black")) + # Darker & Ligher mix aesthetic
  facet_grid(.~label_long, space="free", scale="free") +
  theme_bw() +
  theme(panel.grid=element_blank(),
        axis.text.x=element_blank(),
        legend.position="top",
        strip.text=element_text(size=12))

## Create barplot - proportional barplot of MOI
plot_bar_MOI_fill <- ggplot(file_isolates_MOI_100, aes(x=Isolate, y=as.numeric(MOI), fill=Species)) +
  geom_bar(stat="identity", position="fill") +
  scale_x_discrete(name="Isolate") +
  scale_y_continuous("Proportion of Complexity") +
  scale_fill_manual(name="Species", values=c("#548B54", "#79CDCD", "#8B3E2F", "black")) + # Darker & Ligher mix aesthetic
  facet_grid(.~label_long, space="free", scale="free") +
  theme_bw() +
  theme(panel.grid=element_blank(),
        axis.text.x=element_blank(),
        legend.position="none",
        strip.text=element_text(size=12))

##COMBINE FIGURES
plot_bar_MOI_combine <- ggarrange(plot_bar_MOI_stack, plot_bar_MOI_fill,
                                  nrow=2, align="v", heights=c(1, 0.9), labels=c("a", "b"))

## Figure 4
# Save as image (tiff)
ggsave("Figure4.tiff", plot_bar_MOI_combine,
       device="tiff", dpi=300, width=10, height=7)
# Display
plot_bar_MOI_combine
```

## Figure 5
```{r}
## Download shapefile for Ghana from website: https://gadm.org/download_country.html
# Load the GADM shapefile
ghana_gadm_sf <- st_read("data/gadm41_GHA_shp")

# Check the names of the regions
unique(ghana_gadm_sf$NAME_1)

## Ghana Map
# Combine geometries to get region-level centroids
ghana_regions_combined <- ghana_gadm_sf %>%
  group_by(NAME_1) %>% # Group by region name (adjust column name if different)
  summarize(geometry = st_union(geometry)) # Combine geometries within each region

ghana_centroid_data <- st_centroid(ghana_regions_combined) %>% # Calculate centroids for each combined region geometry
  st_coordinates() %>% # Extract coordinates of the centroids
  as.data.frame() %>%
  mutate(region=ghana_regions_combined$NAME_1) %>% # Adjust column name if needed
  rename(lon=X, lat=Y)

## Bongo Map
# Combine geometries to get district-level centroids for Bongo
bongo_centroid_data <- st_centroid(ghana_gadm_sf) %>% # Calculate centroids for each district
  st_coordinates() %>% # Extract coordinates of the centroids
  as.data.frame() %>%
  mutate(district=ghana_gadm_sf$NAME_2, # Adjust this to match the district column name
         region=ghana_gadm_sf$NAME_1) %>% # Region name, if needed
  rename(lon=X, lat=Y)

## Load prevalence data
MAPs_data_subset <- read_table_in("data/MAPS_prevalence_data.txt") %>%
  select(AgeGrp, Species, Country=`National Unit`, Region=Name, Prevalence=Value, MapType) 

S8MRS_prevalence_data <- read_table_in("data/S8MRS_DBS_Bongo_prevalence_data.txt") %>%
  select(AgeGrp, Species, Country, District, Prevalence, MapType)

##COMBINE DATA
# Join MAPs data to region-level map
MAPs_data_sf <- ghana_gadm_sf %>%
  left_join(MAPs_data_subset, by=c("COUNTRY"="Country", "NAME_1"="Region"))

# Join S8 DBS data to district-level map & add other regions with prevalence=NA
maptypes_list <- c("Survey DBS (2-10yr)", "Survey DBS (all ages)", "Survey DBS (all ages) - adjusted for deep sampling")

bongo_data_sf <- ghana_gadm_sf %>%
  left_join(S8MRS_prevalence_data, by=c("COUNTRY"="Country", "NAME_2"="District")) %>%
  filter(NAME_2 == "Bongo")

non_bongo_data_sf <- map_dfr(maptypes_list, function(sub) {
  MAPs_data_sf %>%
    filter(NAME_2 != "Bongo") %>%
    filter(Species != "(Pm|Po)") %>%
    mutate(MapType=sub) %>%
    mutate(Prevalence=NA)
})

# Combine Bongo data and the synthetic non-Bongo rows
bongo_data_sf_full <- bind_rows(bongo_data_sf, non_bongo_data_sf)

# Combine all map data
merged_data_ALL <- rbind(MAPs_data_sf, bongo_data_sf_full)

## Set factor levels
merged_data_ALL$MapType <- factor(merged_data_ALL$MapType, levels=c("Malaria Atlas (2-10yr)", "Survey DBS (2-10yr)", "Survey DBS (all ages)", "Survey DBS (all ages) - adjusted for deep sampling"))

## Set Color scale
pal <- wes_palette("Zissou1", 100, type = "continuous")

ghana_min <- 0
ghana_max <- merged_data_ALL$Prevalence %>% na.omit() %>% max()

## GHANA MAP
# Plot map - Ghana (Pf only)
bongo_coords <- data.frame(district="Bongo", lat=10.9090167, lon=-0.8079772)

map_ghana <- merged_data_ALL %>%
  filter(MapType=="Malaria Atlas (2-10yr)" & Species=="Pf") %>%
  ggplot() +
  geom_sf(aes(fill=Prevalence), colour="white") +
  geom_text(data=ghana_centroid_data, aes(x=lon, y=lat, label=region), size=3, color="black") +  # Add labels
  geom_label_repel(data=bongo_coords, aes(x=lon, y=lat, label=district), label.padding=0.35, point.padding=0.1, size=3, nudge_y=0.5, nudge_x=0.1) +
  ggtitle(expression("MAP: " * italic("P. falciparum") * " Infection Prevalence by region in Ghana (2-10yr)")) +
  scale_fill_gradientn(name="Prevalence", colours=pal, limits=c(ghana_min, ghana_max)) + 
  theme_pubr() +
  theme(legend.position="right",
        legend.title=element_text(size=11),
        axis.text=element_text(size=11),
        plot.title = element_text(size=12),
        strip.text = element_text(size=12))

## BONGO MAP
# Plot map - Bongo with bordering regions
merged_data_ALL_Bongo <- merged_data_ALL %>%
  mutate(Species=recode(Species, Pf="P. falciparum", Pm="P. malariae", Po="P. ovale spp."),
         Prevalence=ifelse(NAME_2=="Bongo", Prevalence, NA))

map_bongo <- ggplot(merged_data_ALL_Bongo) +
  geom_sf(aes(fill=Prevalence), colour="grey50") +
  coord_sf(xlim=c(-1, -0.6), ylim=c(10.8, 11.01)) +
  geom_text(data=filter(merged_data_ALL_Bongo, NAME_2=="Bongo"), aes(x=-0.7937153, y=10.92294, label=format(Prevalence, digits=3))) +
  scale_fill_gradientn(name="Prevalence", colours=pal, limits=c(ghana_min, ghana_max), na.value="white") + 
  ggtitle(expression(italic("Plasmodium") * " spp. Infection Prevalence (%) in Bongo, Ghana")) +
  facet_grid(MapType~Species, switch="y", labeller=label_wrap_gen(width=16, multi_line=TRUE)) +
  theme_bw() +
  theme(legend.position="none",
        panel.grid=element_blank(),
        axis.text=element_blank(),
        axis.title=element_blank(),
        axis.ticks=element_blank(),
        plot.title = element_text(size=12),
        strip.text.y.left=element_text(angle=0),
        strip.text = element_text(size=12))

## Combine Ghana and Bongo maps
ghana_bongo_combine_maps <- ggarrange(map_ghana, map_bongo, nrow=1, widths=c(2,2.6))
  
## Figure 5
# Save as image (tiff)
ggsave("Figure5.tiff", ghana_bongo_combine_maps,
       device="tiff", dpi=300, width=11, height=6)
# Display
ghana_bongo_combine_maps
```
